---
title: "Final Project GDAT 515"
author: "Emily Arnold"
date: "3/2/2020"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    theme: journal
runtime: shiny
vertical_layout: scroll
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

req.packages <- c("tidyverse", "ggplot2", "plotly", "scales", "lubridate")

'%notin%' <- Negate('%in%')
if(any(installed.packages() %notin% req.packages)){
  missing.packages <- which(req.packages %notin% installed.packages())
  install.packages(req.packages[missing.packages])
}
eval(
  parse(
    text = paste0('library(',req.packages, ')')
  )
)

#polling data was here https://projects.fivethirtyeight.com/polls/president-primary-d/national/

#donation data was here https://www.fec.gov/campaign-finance-data/contributions-individuals-file-description/
```


```{r load data, include=FALSE}

polls <- read.csv("C:/Users/Emily/Documents/Data Science/Data Visualization/final project/earnold-finalproject/president_primary_polls.csv")

polls_select <- polls[,c("candidate_name", "pct", "end_date", "state", "office_type", "party")]

polls_select$candidate_name <- toupper(polls_select$candidate_name)
polls_select$last_name <- str_extract(polls_select$candidate_name, "\\s.*")
polls_select$last_name <- gsub(" ", "", polls_select$last_name)
polls_select$first_name <- str_extract(polls_select$candidate_name, ".+?(?=\\s)")
polls_select$merge_name <- paste0(polls_select$last_name, ", ", polls_select$first_name)
 
polls_select$candidate_name <- NULL
polls_select$first_name <- NULL
polls_select$last_name <- NULL

#need to convert date to actual date
polls_select$end_date <- as.character(polls_select$end_date)
polls_select$end_date <- as.Date(polls_select$end_date, "%m/%d/%y")

polls_select$merge_name[polls_select$merge_name == "R.BIDENJR., JOSEPH"] <- "BIDEN, JOSEPH"
polls_select$merge_name[polls_select$merge_name == "D.HARRIS, KAMALA"] <- "HARRIS, KAMALA"
polls_select$merge_name[polls_select$merge_name == "A.BOOKER, CORY"] <- "BOOKER, CORY"
polls_select$merge_name[polls_select$merge_name == "CASTRO, JULIáN"] <- "CASTRO, JULIAN"
polls_select$merge_name[polls_select$merge_name == "E.GILLIBRAND, KIRSTEN"] <- "GILLIBRAND, KIRSTEN"
polls_select$merge_name[polls_select$merge_name == "F.BENNET, MICHAEL"] <- "BENNET, MICHAEL"
polls_select$merge_name[polls_select$merge_name == "F.WELD, WILLIAM	"] <- "WELD, WILLIAM"
polls_select$merge_name[polls_select$merge_name == "K.DELANEY, JOHN"] <- "DELANEY, JOHN"

polls_select[polls_select$merge_name == "TRUMP, DONALD", "party"] <- "REP"
polls_select[polls_select$merge_name == "DELAFUENTE, ROQUE", "party"] <- "REP"

#only want data for presidential candidates and polls conducted recently
polls_agg <- polls_select %>% 
  filter(.,end_date > "2019-10-01") %>% 
  filter(., office_type == "U.S. President") %>% 
  group_by(merge_name, party) %>% 
  summarize(status = mean(pct))

#reading in campaign contributions data
money <- read.delim("weball20.txt", header = FALSE, sep = "|")

money <- money[,c(2,6,12,18)]
colnames(money) <- c("candidate_name", "total_receipts", "cand_contrib", "individual_contrib" )

money$candidate_name <- gsub("\\.", "", money$candidate_name)
money <- aggregate(.~candidate_name, data = money, FUN = sum)
money$candidate_name <- as.character(money$candidate_name)

#cleaning up data
money$last_name <- str_extract(money$candidate_name, ".+?(?=,)")
money$first_name <- str_extract(money$candidate_name, ",.*")
money$first_name <- gsub(", ", "", money$first_name)
money$first_name <- gsub("\\s.*", "", money$first_name)
money$first_name <- gsub(" ", "", money$first_name)

money$merge_name <- paste0(money$last_name, ", ", money$first_name)

#money$merge_name <- money$candidate_name
money$candidate_name <- NULL
money$last_name <- NULL
money$first_name <-NULL

candidate_data <- na.omit(merge(polls_agg, money, by = "merge_name", all.x = TRUE))

```


Exploring Approval Ratings 
==================================================


This tab explores approval data collected from FiveThirtyEight, and the following charts represent average approval ratings since October 2019. 

column 
-----------------------------------------------------------------------


```{r first visual, echo = FALSE}

plot1_data <- candidate_data %>% 
  top_n(20, status) %>% 
  mutate(merge_name = factor(merge_name, levels = unique(candidate_data$merge_name)[order(candidate_data$status,decreasing =TRUE)]))

plot_ly(data = plot1_data, x = ~status, y = ~merge_name, 
        text = ~paste0("Name: ", merge_name, "- ", party)) %>% 
        add_trace(p = ., type = "bar") %>% 
        layout(xaxis = list(title = "Polling Average", showgrid = FALSE),
         yaxis = list(title = "Candidate", showgrid = FALSE),
         title = "Overall Approval Rating Average by Candidate")

```


column 
-----------------------------------------------------------------------

```{r third chart, echo=FALSE}

names <- candidate_data$merge_name

limited_over_time <- polls_select %>% 
                    filter(merge_name %in% names) %>% 
                    select(c(pct, end_date, merge_name)) 


limited2 <- limited_over_time
limited2$end_date = format(as.Date(limited2$end_date),"%Y-%m")
limited2 <- as.data.frame(aggregate(.~end_date+merge_name, data = limited2, FUN = mean))

test <-limited2 %>% 
  top_n(2000, pct) %>% 
ggplot(., aes(x = end_date, y = pct, color = merge_name))+
  geom_point(aes(color = merge_name))+
  labs(title = "Approval Ratings Over Time",
       x = "Date",
       y = "Approval rating average")+
    theme_minimal()+
    theme(legend.position = "none", axis.text.x = element_text(angle = 45)) 

ggplotly(test)

```


Exploring Campaign Finances
==================================================

This data is from the FEC, and represents total campaign funding, as well as individual contributions.

column
-----------------------------------------------------------------------

```{r money only chart, echo = FALSE}

candidate_data %>%
  top_n(20, total_receipts) %>%
  arrange(.,desc(total_receipts)) %>%
  mutate(total_receipts = total_receipts / 10^6) %>%
  mutate(cand_contrib  = cand_contrib / 10^6) %>%
ggplot(data = ., aes(x = reorder(merge_name, (total_receipts)), y = total_receipts))+
  geom_col(aes(fill = cand_contrib ))+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Donations", x = "Candidate", y = "Campaign Money")+
  scale_fill_gradient(low = "#5da5b6", high = "#FF6347", name = "Personal Contribution (M)")+
  scale_y_continuous(labels = unit_format(unit = "M"))+
  theme_minimal()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  coord_flip()

```




Deep Dive
==================================================

Select a candidate to reveal information about finances and polling results.

```{r selection bar, echo=FALSE}

candidates <- candidate_data$merge_name

selectInput("merge_name", label = "Choose a Candidate:",
            choices = candidates, selected = candidates[1])

```

The following chart shows campaign funding vs. approval ratings. There is generally a positive correlation between the two.

column
-----------------------------------------------------------------------

```{r echo = FALSE}

candidate_data2 <- candidate_data %>%
  mutate(total_receipts = total_receipts / 10^6)

renderPlot({
  ggplot(data = candidate_data2, aes(x = total_receipts, y = status))+
  geom_point()+
  geom_point(data = candidate_data2[candidate_data2$merge_name == input$merge_name,],
             aes(x = total_receipts, y = status),
             colour = "red",
             size = 5)+
  scale_x_continuous(labels = unit_format(unit = "M"))+
  labs(title = "Status vs. Donations - All Candidates",
       x = "Campaign Money",
       y = "Approval rating average")+
    theme_minimal()
})

```

This chart displays the selected candidates approval over time.

```{r plot for candidate polling over time, echo = FALSE}

renderPlot({
  
  limited_over_time %>%
  filter(merge_name == input$merge_name) %>% 
  filter(.,end_date > "2019-10-01") %>% 
  ggplot(data = ., aes(x = end_date, y = pct))+
  geom_point()+
  geom_smooth(colour = "#00688B", se = FALSE)+
  labs(title = "Approval Over Time", 
       x = "Date", 
       y = "Approval rating average")+
    theme_minimal()
})


```

